### 2014-03-15 src/analysis.Rnw
All Alex edits

\section{Analysis}
\label{sec:analysis}
For the purposes of comparing between the criteria, we computed the
correlation between sampled mutation score and full mutation score.
Since for most of the sampling criteria, if there are no mutants to
sample, it probably means there are no mutants generated on the whole,
we used
\[\mu_\{M_{all}|M_{reduced}\} = 0 + \beta_1 \times M_{reduced}\]
as the regression formula, and reported
the correlation for this linear regression. We also report the expected
multiplier factor $\beta_1$ for the sampled mutation score. This factor
determines the ease of satisfying the sampling criteria. That is, if the
multiplier factor is small, then it is easier to satisfy compared to a
larger multiplier factor.

Next, for each criteria, we computed the reduction of mutants given by the
ratio between total number of mutants and sampled mutants.
We also compute standard deviation of the mutant reduction, which determines
the consistency of results. A large amount of standard deviation suggests
that the mutant reduction ratio achieved is very much dependent on the
project. So as a heuristic, we suggest looking at only those methods which
have a mean reduction ratio at least twice the standard deviation.

A few projects did not return any results for some large reduction factors.
Thus number of valid results from projects is also collected in the result
tables.

>>>>

\section{Analysis}
\label{sec:analysis}
For the purposes of comparing between the criteria, we computed the
linear correlation between sampled mutation score and full mutation score.
Since for most of the sampling criteria, if there are no mutants to
sample, it probably means there are no mutants generated,
we used
\[\mu_\{M_{all}|M_{reduced}\} = 0 + \beta_1 \times M_{reduced}\]
as the regression formula, and reported
the correlation for this linear regression. We also report the expected
multiplier factor $\beta_1$ for the sampled mutation score. This factor
determines the ease of satisfying the sampling criteria. That is, if the
multiplier factor is small, then it is easier to satisfy compared to a
larger multiplier factor.

Next, for each criteria, we computed the reduction of mutants given by the
ratio between total number of mutants and sampled mutants.
We also compute standard deviation of the mutant reduction, which determines
the consistency of results. A large standard deviation suggests
that the mutant reduction ratio achieved is very much dependent on the
project. Therefore, as a heuristic, we suggest looking at only those methods which
have a mean reduction ratio at least twice the standard deviation.

A few projects did not return any results for some large reduction factors.
Thus the number of valid results from projects is also collected in the result
tables.

### 2014-03-15 src/discussion.Rnw
update
All Alex edits

From our data, we see that for original and generated test suites, the
mutant reduction ratio and standard deviation was similar for stable strategies
(where the standard deviation of the reduction achieved was less than half
of the reduction achieved). This suggests that our analysis is equally
applicable for both generated test suites and original test suites.

Analysing the scores achieved further, we note that the accuracy of prediction
($R^2$) achieved by sampling is generally lower for generated test
suites. Especially, as the fraction of sampling becomes larger the accuracy
reduces drastically for generated test suites while the accuracy still
holds good for original test suites
Secondly, we also observe that the multiplier factor $\beta_1$ 
is smaller for generated test suites than original test suites. This
observation holds good for all sampling and operator selection strategies
examined.  Both these factors together suggest that generated
test suites have a higher variability in their sampled mutation scores than
original test suites. We note that since $\beta_1$ was generally smaller
for generated test suites than original test suites, the full mutation score
corresponding to a sampled mutation score for generated test suites would
be lower than what would be expected from original test suite.
Secondly, the amount of reduction that can be gained by sampling satisfying
a given accuracy is lower for generated test suites in comparison with original
test suites.

Considering different sampling strategies, we see a few general patterns. The
first is that sampling strategy can predict the final score with very
minimal loss in accuracy up to a fraction of $\frac{1}{64}$ of the total
number of mutants (Our experiments suggest up to $70$ times reduction,
with a low standard deviation of $6.72$). We did not cut off with the 99\%
rule as mentioned by Zhang et al.~\cite{zhang2013ase} (Originally proposed
by Offutt et al.~\cite{offutt1996experimental} for a different measurement),
and instead chose to observe that we obtained $R^2 > 0.96$ which provides
a high confidence in the accuracy of prediction.

The second pattern is that fractional sampling based on projects appears
to be better than that based on classes, which in turn is better than
methods and lines, with regard to the stability of the result (i.e. low
standard deviation).  This seems not very visible at lower levels (where
the reduction fraction is less than 32), but becomes more pronounced as the
mutant reduction fraction increases.  There seem to exist a hierarchy of
sampling, where by as sampling becomes more fine grained stability decreases.
We did not see the pattern found by Zhang et al.~\cite{zhang2013ase} where
method-based sampling had the highest accuracy. In fact, in our measurements,
the opposite may be true, with project based sampling getting the highest
correlation with the final score a majority of the time.

The third pattern is that the simplest sampling scheme --- simple random sampling ---
performs best compared to sampling schemes that tried to provide higher weight
to specific elements based on specific criteria. More importantly, operator selection
seems to be generally worse off in predicting the final score than any of the random
selection schemes.

>>>>

% About Randoop and Original
From our data, we see that for original and generated test suites, the
mutant reduction ratio and standard deviation were similar for stable strategies
(where the standard deviation of the reduction achieved was less than half
of the reduction achieved). This suggests that our analysis is equally
applicable for both original test suites and generated ones.

% About Randoop and Original
Analysing the scores further, we note that the accuracy of prediction
($R^2$) achieved by sampling is generally lower for generated test
suites. As the sampling denominator increases, the accuracy
reduces drastically for generated test suites while correlation remains strong
for original test suites
Secondly, we also observe that the multiplier factor $\beta_1$ 
is smaller for generated test suites than original test suites. This
observation holds good for all sampling and operator selection strategies
examined.  Both these factors together suggest that generated
test suites have a higher variability in their sampled mutation scores than
original test suites. We note that since $\beta_1$ was generally smaller
for generated test suites than original test suites, the full mutation score
corresponding to a sampled mutation score for generated test suites would
be lower than what would be expected from the original test suite.
Secondly, the amount of reduction that can be gained by sampling that must reach a given correlation is lower for generated test suites.

% Different sampling strategies
Considering different sampling strategies, based on their stability and mutation
reduction achieved, we note that using simple sampling (\textit{sample(x)})
produced the most stable results. For both generated and original test suites,
this strategy produced valid results for all different scopes and sampling
ratios. It also was able to achieve the highest mutation reduction of $72.63$,
that is only $1.37\%$ of the original mutants were needed to produce a highly
correlated ($R^2 = 0.97$) mutation score, with minimal deviation
($\sigma = 6.36$) for original test suites. While the correlation is weaker for
generated test suites it is still quite strong, with $R^2 = 0.90$

Our \textit{round(x)}, the strategy of ignoring low complexity elements,
did not pay off, with line scope having a tendency to become unstable and
inaccurate at high sampling ratios. At very low sampling ratio ($1/64$), the
strategy with method scope also becomes unstable and inaccurate. On the whole
\textit{round(x)} did not perform as well as \textit{sample(x)}

The strategy of \textit{ceil(x)} also did not perform well. While the
results were generally stable, the reduction ratios achieved were lower than
\textit{sample(x)}, with similar correlations.

For line count strategies, only the external sampling produced
consistent results, which were again not as good as \textit{round(x)}
for similar mutation reduction ratios. Secondly, we had hypothesised
that perhaps coverage was a significant contributor to the accuracy of
sampled mutation score. However we could find no evidence in support
of this hypothesis with line, method, class, and project scope showing
similar accuracy and consistency at similar reduction ratios.

The one per element strategy did not have stable results in general except
for at the scope of line (at which point it was same as the count-per-element
strategy).

While the x\% per operator strategy produced consistent results, its
consistency was not as good as \textit{sample(x)}, and was either on
par with or worse than \textit{sample(x)} in correlation.

Considering mutation operators, only three operators \textit{NC},
\textit{RC},and \textit{DC} had consistent results (and only for
original test suites). The n-selection strategy had high accuracy and
consistency (though low reduction ratio) until 7-selection (removing
the seven most numerous operators) for both generated and original
test suites.  Javalanche did not have a high mutation reduction ratio,
though it had high correlation for both kinds of test suites. The other
operator selection methods performed poorly on generated test suites,
while they had high correlation and reasonable stability on original test
suites. However, for comparable mutation reduction levels, they all performed
worse than random sampling strategies.

Considering different sampling strategies, we see a few general patterns. The
first is that sampling strategy can predict the final score with very
minimal loss up to a fraction of $\frac{1}{64}$ of the total
number of mutants (our experiments suggest up to $70$ times reduction,
with a low standard deviation of $6.72$). We did not cut off with the 99\%
rule as mentioned by Zhang et al.~\cite{zhang2013ase} (originally proposed
by Offutt et al.~\cite{offutt1996experimental} for a different measurement),
and instead chose to observe that we obtained $R^2 > 0.96$ which provides
a high confidence in the accuracy of prediction.

The second pattern is that fractional sampling based on projects
appears to be better than that based on classes, which in turn is
better than methods and lines, with regard to the stability of the
result (i.e. low standard deviation).  This is not clear at lower
levels of reduction (more than 1/32), but becomes more pronounced as
the denominator of the mutant reduction fraction increases.  There
seems to exist a hierarchy of sampling, whereby as sampling becomes
more fine grained stability decreases.  We did not observe the pattern
found by Zhang et al.~\cite{zhang2013ase} where method-based sampling
had the best correlations. In fact, in our measurements, the opposite
may be true, with project based sampling having the highest
correlation with the total mutation score a majority of the time.

The third "pattern" is that the simplest sampling scheme --- simple
random sampling --- performs best compared to sampling schemes that
tried to provide higher weight to specific elements based on specific
criteria. Additionally, operator selection seems to be generally worse
off in predicting the final mutation score than any of the random selection
schemes.  In principle, if we could devise sampling methods that
somehow captured only the most discriminative mutants for programs,
they might perform better than any of these; our results suggest that
most schemes that (we assume) aim at this goal do not in fact achieve
the goal over a large body of programs, and so tend to perform worse
than simply randomly sampling the whole pool of mutants, which has
attractive simplicity in the absence of a better strategy.

### 2014-03-15 src/introduction.Rnw
All Alex edits

Recent work~\cite{zhang2010icse,zhang2013ase} has investigated the
relative merits of random sampling strategies and operator selection.
Random sampling can perform as well as or better than operator
selection, in these studies; a strategy of either program elements
based sample or one combining both program element-based sampling
and operator selection was best.

However, as pointed out by Zhang et al.~\cite{zhang2013ase}, the field
has a serious lacuna in large scale research, both in the size of the
programs studied, and in the number and diversity of programs, which
reduces our confidence in all results. This is true for selective
mutation studies, studies in sampling, and also for comparatively
newer studies that attempt to combine methods. This is particularly
worrisome if mutation analysis is to gain wider acceptance among
testing professionals. Further, quite a few of the influential
studies~\cite{offutt1993experimental,offutt1996anexperimental,wong1995reducing,mresa1999efficiency}
were conducted on older programming languages such as Fortran, with
operators specific to the language, and are not directly applicable to
newer languages such as Java. Finally, modern research into the byte
code based mutation (a \textit{do faster} approach for eliminating
the compilation step to gain execution speed) operators based on
source code modification are no longer applicable, and their
equivalents in bytecode need to be identified and compared with other
approaches. A detailed discussion of these issues can be found in
Section~\ref{sec:related}.

We have attempted to rectify this situation with a large scale study
of real world programs.  As detailed in Section~\ref{sec:methodology},
we sample 188 Java programs from Github, with size ranging from 50 LOC
to 100 KLOC, which allows widely applicable statistical inferences to
be made. We provide the results from running both original project
human generated test suites and also automatically generated
test suites using random testing, increasing the applicability of our
findings to both testers in traditional development who are interested in applying mutation analysis to suite evaluation and to researchers who seek to compare test suites quickly.

@@

\item Our recommendations are valid for both testing researchers who
might be comparing automatically generated test suites, and software
testers in real world development working to produce and evaluate manual test suites.

\item Most importantly, we find that random sampling strategies perform
better than operator selection strategies consistently in predicting the
final mutation score.

>>>>

Recent work~\cite{zhang2010icse,zhang2013ase} has investigated the
relative merits of random sampling strategies and operator selection.
Random sampling can perform as well as or better than operator
selection, in these studies; a strategy of either sampling based on
program elements or one combining both program element-based sampling
and operator selection was best.

However, as pointed out by Zhang et al.~\cite{zhang2013ase}, the field
has a serious lacuna in large scale research, both in the size of the
programs studied, and in the number and diversity of programs, which
reduces our confidence in all results. This is true for selective
mutation studies, studies in sampling, and also for comparatively
newer studies that attempt to combine methods. This is particularly
worrisome if mutation analysis is to gain wider acceptance among
testing professionals. Further, quite a few of the influential
studies~\cite{offutt1993experimental,offutt1996anexperimental,wong1995reducing,mresa1999efficiency}
were conducted on older programming languages such as Fortran, with
operators specific to the language, and are not directly applicable to
newer languages such as Java. Finally, due to modern research into the byte
code based mutation (a \textit{do faster} approach for eliminating
the compilation step to gain execution speed), operators based on
source code modification are no longer applicable, and their
equivalents in bytecode need to be identified and compared with other
approaches. A detailed discussion of these issues can be found in
Section~\ref{sec:related}.

We have attempted to rectify this situation with a large scale study
of real world programs.  As detailed in Section~\ref{sec:methodology},
we sample 188 Java programs from Github, with size ranging from 50 LOC
to 100 KLOC, which allows widely applicable statistical inferences to
be made. We provide the results from running both the original
project's (presumably human-generated) test suites and also
automatically generated test suites using random testing, increasing
the applicability of our findings to both testers in traditional
development who are interested in applying mutation analysis to suite
evaluation and to researchers who seek to compare test suites quickly to evaluate testing techniques.

One note about our, and other researchers, evaluation criteria is in
order.  We, like others, use the ability of a sample to correlate well
to the score over all mutants as a measure of sampling quality.  The
underlying reason for this is that, while many mutants have little
ability to distinguish between good and bad test suites, some smaller
set of mutants show subtle deficiencies in suites that are not clear
from, say, code coverage.  The full set of mutants obviously contains
these mutants, and correlating with (rather than exactly matching) the
full mutation score is a strong indicator that many of these
discriminatory mutants have been preserved in a sample.

@@

\item Our recommendations are useful for both testing researchers who
might be comparing automatically generated test suites, and software
testers in real world development working to produce and evaluate manual test suites.

\item Most importantly, we find that random sampling strategies perform
better than operator selection strategies in consistently predicting the
final mutation score while obtaining stable computational reduction.

### 2014-03-15 src/methodology.Rnw
Add x\% selection for operators
Fix citation for wong
Fix space requirements
All Alex edits

\section{Methodology} \label{sec:methodology} Our selection of
programs was driven by a few overriding concerns. Our primary
requirement was that our results had to be as widely applicable as
possible for real-world programs. Secondly, we strived for a
statistically significant result, reducing the number of variables
present in the experiments. For this reason we chose a random sample
of Java projects from Github\cite{github} that use the popular
maven\cite{maven} build system, following the methodology used by some
of the authors in an ICSE 2014 paper~\cite{gopinath2014code}. As in
that work, we utilize two kinds of of test suites: the original test
suites present in a limited number of projects, and automatically
generated test suites from Randoop~\cite{randoop}.  Due to limited
resources, we had to further reduce the programs and test suites to those
that could finish mutation analysis in 10 minutes. From this set of test
suites, we sampled 110 suites each for both original and Randoop generated
test suites. Since there was some overlap between projects with original
test suites, and those with Randoop test suites, we had 81 unique projects
each in both sets. The remaining projects had both flavors of test suites.
% {\bf ADG I have no idea what that last sentence means, really.
% RG: Fixed}

@@

\subsubsection{Lines per element and variants}
Our previous research~\cite{gopinath2014code} found that statement coverage
was highly correlated with mutation coverage for a project. This immediately
suggests that perhaps choosing one mutant per line may be sufficient to achieve
a close approximation of the final mutation score. Further, statement deletion
has been researched previously~\cite{untch2009onreduced,deng2013empirical} and
has been found to reduce the number of mutants well, without an appreciable
decrease in accuracy. This also provides a nice comparison with similar
numbers of mutants between operator selection and random sampling of mutants.

@@

For the first variant, we applied the x\% sampling to the result of the first
sampling based on line counts. That is, in the case of methods, we first
selected \textit{linecount} mutants each from each method, and applied the x\%
selection to the result of this selection. For e.g for $1/2$ sampling on methods
if a method had $10$ lines and another had $10$ lines, we sampled $10$ mutants
from the first, and $20$ mutants from the second, and from the combined $30$
mutants, we sampled $15$.  For the second variant, we did the sampling in one go,
where we chose \textit{linecount/x} number of mutants from each program element.
That is, in the previous example, only $5$ mutants from the first and $10$ mutants
from the second method would be chosen.

@@

\subsubsection{One per element}
Another simple strategy of sampling we experimented with is to just choose
one mutant per program element, in the order \textit{line},\textit{method},
\textit{class}.

@@

For selective methods, we tried mutation operators suggested by Wong et al.~\cite{wong1995reducing}, Offutt et al.~\cite{offutt1996anexperimental,deng2013empirical}, % TODO Barbosa
 and, Namin et al~\cite{siami2008sufficient}. Since Javalanche~\cite{javalanche} utilized operator selection mechanisms, we also compared the Javalanche operators for operator selection. Note that all of these techniques except 
Javalanche have targeted C programs, thus some of these operators
are sensible in C but not in Java programs. For example C compilers tolerate
removing a return statement --- in Java this causes a compilation error.
Moreover, there were some operators that are not supported by the PIT framework
like ABS (absolute mutation operators). We tried to realize as many
as possible of the operators using PIT's bytecode level mutation. We also extended 
the PIT framework to support the  statement deletion operator (SDL).

@@

\subsubsection{Criteria I}
For this criteria, we used the two operators \pitcb, and \pitnc which are similar to those given by Wong. We did not have ABS operator available in PIT.
\subsubsection{Criteria II}
We used the operators given by Offutt et al. \pitin, \pitm, \pitcb, \pitnc. We did not have LCR available in PIT.
\subsubsection{Criteria III}
Those given by Javalanche \pitnc, \pitvmc, \pitnvmc, \pitm, \pitic, \pitemv. 
\subsubsection{Criteria IV}
By Namin et al. \pitin, \pitm, \piti, \pitnc. Many of operators were not applicable in Java and some were not available in PIT like IndVarBitNeg.
\subsubsection{Criteria V}
For this criteria, we utilized the approach suggested by
Offutt et al.~\cite{offutt1993experimental} where by most numerous operators
are discarded at each step.

\subsubsection{Statement Deletion:}
\label{subsec:stmtdel}
The basic statement deletion was modeled on the work by Deng et
al.~\cite{deng2013empirical}. The operations on single statements was modeled
using \pitvmc, \pitnvmc, \pitcc, \pitemv, \pitri
for simple statements, and using \pitrc for control structures,
which replaced the boolean condition with false, resulting in removal of
the block of statements. The operator for return values was modeled using
\pitrv, which is similar.  The operators for \textit{while,for},
and \textit{if} statements were modeled using \pitdc,
which replaced the boolean condition with true, which removed the effect
of conditional. The \textit{switch} statement deletion was modeled using
\pitrs which replaced the first 100 labels with default label,
resulting in the switch element getting deleted. (Due to the constraints
of the architecture of PIT only the first 100 labels were replaced). The
\textit{try catch} was not necessary at bytecode level.  Finally, to get an
accurate estimation of the effect of true statement deletion, we grouped
the mutants by line, and considered a single virtual mutant (That is, the
total number of mutants is equal to the total number of lines). Further,
killing any mutant from a line resulted in marking the virtual mutant
for that line as killed. This gave us the mutation score for the virtual
statement deletion operator.  We note that the approximation of simple
statement deletion, especially when arithmetic operators are involved is not
complete. However, as seen in Figure~\ref{fig:op}, the number of mutants
produced by \pitm operator is very small. We also note that our
approximation does not account for the increase in ease of detection when
multiple mutations are combined together due to coupling effect. This also
means that not all the lines may be mutated since there may be no applicable
operators. However, given the constraints of a bytecode based mutation system,
we believe that our procedure is reasonable. An interesting point is that
by this procedure statement deletion is no longer a simple mutation and qualifies
as a higher order mutation operator. In fact, the procedure we outlined can
be used in general to judge the effectiveness of a higher order operator from
the results of first order mutation operators that makeup the higher order
mutant in question.

>>>>

\section{Methodology} \label{sec:methodology} Our selection of
programs was driven by a few overriding concerns. Our primary
requirement was that our results had to be as widely applicable as
possible for real-world programs. Secondly, we strived for a
statistically significant result, reducing the number of variables
present in the experiments. For this reason we chose a random sample
of Java projects from Github~\cite{github} that use the popular
maven~\cite{maven} build system, following the methodology we used
previously in an ICSE 2014 paper~\cite{gopinath2014code}. As in that
work, we utilize two kinds of of test suites: the original test suites
present in a limited number of projects, and automatically generated
test suites from Randoop~\cite{randoop}.  Due to limited resources, we
had to further reduce the programs and test suites to those that could
finish mutation analysis in 10 minutes. From this set of test suites,
we sampled 110 suites each for both original and Randoop generated
test suites. Since there was some overlap between projects with
original test suites, and those with Randoop test suites, we had 81
unique projects each in both sets. The remaining projects had both
flavors of test suites.  % {\bf ADG I have no idea what that last
sentence means, really.  % RG: Fixed}

@@

\subsubsection{Lines per element and variants}
Our previous research~\cite{gopinath2014code} found that statement coverage
was highly correlated with mutation score for a project. This immediately
suggests that perhaps choosing one mutant per line may be sufficient to achieve
a close approximation of the final mutation score. Further, statement deletion
has been researched previously~\cite{untch2009onreduced,deng2013empirical} and
has been found to reduce the number of mutants well, without an appreciable
decrease in accuracy. This also provides a nice comparison with similar
numbers of mutants between operator selection and random sampling of mutants.

@@

For the first variant, we applied the x\% sampling to the result of
the first sampling based on line counts. That is, in the case of
methods, we first selected \textit{linecount} mutants each from each
method, and applied the x\% selection to the result of this
selection. For example, for $1/2$ sampling on methods if one method had
$10$ lines and another had $20$ lines, we sampled $10$ mutants from
the first, and $20$ mutants from the second, and from the combined
$30$ mutants, we sampled $15$.  For the second variant, we did the
sampling in one go, where we chose \textit{linecount/x} number of
mutants from each program element.  That is, in the previous example,
only $5$ mutants from the first and $10$ mutants from the second
method would be chosen.

@@

\subsubsection{One per element}
Another simple strategy of sampling we experimented with is to just choose
one mutant per program element, in the order \textit{line}, \textit{method}, \textit{class}.

\subsubsection{x\% selection per operator}
This strategy, first suggested by Wong et al.~\cite{wong1995reducing} samples an equal percentage
of mutants from each operator.

@@

For selective methods, we tried mutation operators suggested by Wong et al.~\cite{wong1995reducing}, Offutt et al.~\cite{offutt1996anexperimental,deng2013empirical}, % TODO Barbosa
 and, Namin et al~\cite{siami2008sufficient}. Since Javalanche~\cite{javalanche} utilized operator selection mechanisms, we also compared the Javalanche operators for operator selection. Note that all of these techniques except 
Javalanche have targeted C programs, thus some of these operators
are sensible in C but not in Java programs. For example C compilers tolerate
removing a return statement --- but in Java this causes a compilation error.
Moreover, there were some operators that are not supported by the PIT framework
like ABS (absolute mutation operators). We tried to realize as many
as possible of the operators using PIT's bytecode level mutation. We also extended 
the PIT framework to support the  statement deletion operator (SDL).

@@

\subsubsection{Criteria I}
For this criteria, we used the two operators \pitcb and \pitnc which are similar to those given by Wong. We did not have ABS operator available in PIT.
\subsubsection{Criteria II}
We used the operators given by Offutt et al. \pitin, \pitm, \pitcb, \pitnc. We did not have LCR available in PIT.
\subsubsection{Criteria III}
Those given by Javalanche \pitnc, \pitvmc, \pitnvmc, \pitm, \pitic, \pitemv. 
\subsubsection{Criteria IV}
Namin et al. used \pitin, \pitm, \piti, \pitnc. Many of these operators were not applicable in Java, and some were not available in PIT (e.g. IndVarBitNeg).
\subsubsection{Criteria V}
For this criteria, we utilized the approach suggested by
Offutt et al.~\cite{offutt1993experimental} where the most numerous operators
are discarded at each step.

\subsubsection{Statement Deletion:} \label{subsec:stmtdel} The basic
statement deletion was modeled on the work by Deng et
al.~\cite{deng2013empirical}. The operations on single statements were
modeled using \pitvmc, \pitnvmc, \pitcc, \pitemv, and \pitri for
simple statements, and using \pitrc for control structures, which
replaced the boolean condition with false, resulting in removal of the
block of statements. The operator for return values was modeled using
\pitrv, which is similar.  The operators for \textit{while},
\textit{for}, and \textit{if} statements were modeled using \pitdc,
which replaced the boolean condition with true, which removed the
effect of conditional. The \textit{switch} statement deletion was
modeled using \pitrs which replaced the first 100 labels with a
\textit{default} label, resulting in the switch element being
deleted. Due to the constraints of the architecture of PIT only the
first 100 labels were replaced. Deleting \textit{try/catch} was not
necessary at bytecode level.  Finally, to get an accurate estimation
of the effect of true statement deletion, we grouped the mutants by
line, and considered a single virtual mutant (that is, the total
number of mutants is equal to the total number of lines). Further,
killing any mutant from a line resulted in marking the virtual mutant
for that line as killed. This gave us the mutation score for the
virtual statement deletion operator.  We note that the approximation
of simple statement deletion, especially when arithmetic operators are
involved, is not complete. However, as seen in Figure~\ref{fig:op},
the number of mutants produced by \pitm operator is very small. We
also note that our approximation does not account for the increase in
ease of detection when multiple mutations are combined together due to
the coupling effect. This also means that not all the lines may be
mutated ,since there may be no applicable operators. However, given
the constraints of a bytecode based mutation system, we believe that
our procedure is reasonable. An interesting point is that using this
procedure statement deletion is no longer a simple mutation and
qualifies as a higher order mutation operator. In fact, the procedure
we outlined can be used in general to judge the effectiveness of a
higher order operator from the results of the first order mutation
operators that together compose the higher order mutant.

### 2014-03-15 src/results.Rnw
Fix space requirements
All Alex edits

Our results for random sampling over program elements using $sample(x)$
are given in Table~\ref{tbl:xselsample}. The result of the variation using
$round(x)$ is given in Table~\ref{tbl:xselround} and that using
$ceil(x)$ is given in Table~\ref{tbl:xselceil}.
The first column in each table provides the scope of sampling i.e. if it was
done per class, per method, per line etc. The second column provides the
fraction involved.  The third column contains the $R^2$ value obtained between
the mutation score of the sampling criteria used and the full mutation score.
The fourth column contains the mutant reduction factor. This is the average of
total mutants produced divided by the number of mutants sampled by the criteria.
The fifth column contains the standard deviation of the same. A few of the projects
did not return any value for certain sampling criteria for particular mutation
reduction fractions. The sixth column contains the number of projects that
actually had valid values for the criteria being considered. The seventh column
is the average multiplier between the sampled mutation score and the full mutation
score.

Table~\ref{tbl:lcountl} provides the results for line count criteria where the
sampling was conducted outside program element, and Table~\ref{tbl:lcountw} contains
the results for line count criteria where the sampling was conducted inside the
program element. Similarly Table~\ref{tbl:onesel} contains the results for
the criteria of one mutant per program element.

The same data (only for original test suite) is visualized in Figure~\ref{fig:xroundsel}.

Finally Table~\ref{tbl:operators} contains the results when only a single operator
is used, and Table~\ref{tbl:opsel} contains the results for specific operator
selection criteria. The first column in these two tables contain the particular
operator or operator selection applied. Second column contains the
correlation coefficient $R^2$ with the final mutation score. Third column contains
the mean mutant reduction ratio, and fourth contains the standard deviation of the
mutant reduction ratio achieved. The fifth column contains the number of projects
that returned valid values for that operator selection mechanism, and the final
column gives the multiplier factor.
The TSDL row is the result of applying only the higher order mutant
equivalent to \textit{statement deletion}.
All the \textit{N-sel} rows are the results of N-selective operator selection
strategies~\cite{offutt1993anexperimental}. We used the order of operators
determined from Figure~\ref{fig:op} for this purpose.

>>>>

Our results for random sampling over program elements using
$sample(x)$ are given in Table~\ref{tbl:xselsample}. The result of the
variation using $round(x)$ is given in Table~\ref{tbl:xselround} and
that using $ceil(x)$ is given in Table~\ref{tbl:xselceil}.  The first
column in each table provides the scope of sampling: per class, per
method, or per line. The second column provides the fraction involved.
The third column contains the $R^2$ value obtained between the
mutation score of the sampling criteria used and the full mutation
score.  The fourth column contains the mutant reduction factor. This
is the average of total mutants produced divided by the number of
mutants sampled by the criteria.  The fifth column contains the
standard deviation of the same. A few of the projects did not return
any value for certain sampling criteria for particular mutation
reduction fractions. The sixth column contains the number of projects
that actually had valid values for the criteria being considered. The
seventh column is the average multiplier between the sampled mutation
score and the full mutation score.

Table~\ref{tbl:lcountl} provides the results for line count criteria
where the sampling was conducted after combining program elements, and
Table~\ref{tbl:lcountw} contains the results for line count criteria
where the sampling was conducted inside each program
element. Similarly Table~\ref{tbl:onesel} contains the results for the
criteria of one mutant per program element.

Finally, Table~\ref{tbl:operators} contains the results when only a single operator
is used, and Table~\ref{tbl:opsel} contains the results for specific operator
selection criteria. The first columns in these two tables contain the particular
operator or operator selection applied. The second column contains the
correlation coefficient $R^2$ with the final mutation score. The third column contains
the mean mutant reduction ratio, and the fourth contains the standard deviation of the
mutant reduction ratio achieved. The fifth column contains the number of projects
that returned valid values for that operator selection mechanism, and the final
column gives the multiplier factor.
The TSDL row is the result of applying only the higher order mutant
equivalent to \textit{statement deletion}.
All the \textit{N-sel} rows are the results of N-selective operator selection
strategies~\cite{offutt1993anexperimental}. We used the order of operators
determined from Figure~\ref{fig:op} for this purpose. The complete data
(only for original test suites) is visualized in Figure~\ref{fig:fullimgsize}.
